Radio Emission from Red-Giant Hot Jupiters

### 2015-11-12 reply2referee.txt

1) response to ``The authors should demonstrate that the ram pressure of any exoplanetary outflow induced by heating can also be neglected in determining the size of the magnetosphere. ''

2) argue the necesity of ``Observability`` section, in particular S4.3 and S4.4.

3) response to ``it would be useful if they included a brief calculation or statement of how many might be expected, particularly in the age ranges they consider. ''

4) literature survey if 4.3 and 4.4 have been discussed already in previous papers

@@

                                                10 November 2015

Dear Dr. Burton,

  We now resubmit our paper titled "Radio Emission from Red Giant Hot
Jupiters" by Yuka Fujii and collaborators, having addressed the
referee's points.  In so doing we have provided additional explanatory
text and updated the references, both in ways suggested by the referee
and on our own initiative.  Below is our point-by-point response to
the referee.  In our response, we present the referee's comments
preceded by "## ", and our own comments indented in the space that
follows. We also revised the reference list to match the journal's
nomenclature.

  We realized upon reading the referee's report that we should have
done a better job of motivating some of the sections in our paper.
For instance, we wrote the section about the canceling field generated
by stellar-wind electrons, as they spiral down field lines in
accreting onto the planet, in order to allay the reader's possible
concerns that the planetary magnetic field would be significantly
attenuated by the canceling field generated by the massive stellar
wind from an AGB star -- a possibility that we wrestled with for
probably 100 hours.  We hope to save our readers from having to repeat
the same thought process we went through by laying out the logic
clearly.  However, it seems that we didn't explain well enough why we
put this section in the paper; in this resubmission, we have tried to
better motivate the parts that the referee thought were less germane.

@@

  In addition, it appears that the referee and we have somewhat
different philosophies regarding the purpose of a paper.  For example,
the referee suggests that we eliminate the section "Assumptions for
planetary magnetic field" and instead simply state that we assume a
magnetic field of 10G.  In our view, this would significantly weaken
our manuscipt.  Our aim is not merely to explain the assumptions that
go into making each figure and table but also to walk the reader
through the physics of the processes under discussion.

  Finally, we point out that this comment from the report was
surprising:
----
  "Section 2.5 This section is new, but not necessary, as Ignace
   (2010) have already addressed the issue of the ionisation of the
   exoplanetary exosphere."
----
Our manuscript presents a more detailed and accurate description of
the ionisation profile around a planet accreting stellar wind than
that in Ignace et al.'s paper.  It is unusual, in our experience, for
a referee to say that a description that is new, correct, and more
detailed than another description in the literature, is unnecessary.

@@

## Fig 3: the numbers on the RHS have no units...they are probably
## fluxes, but they should be shown.

  We appreciate this feedback. We have indicated the unit in the title
  of Figure 3.

## Section 4:
## 
## Much of this section could be cut, saving perhaps section
## 4.2. Predictions for observations with different facilities have been
## made many times before.

  We believe that Section 4.1 (regarding intrinsic radio emission from
  red giant stars) enhances the paper.  For instance, when we discuss
  this work with people, they often ask us about this topic.  It's a
  natural consideration, since a key component of observability is
  SNR, and the star obviously contributes some of the Noise (indeed,
  in most exoplanet contexts, the signal from the planet is competing
  against a much brighter star).  To devote 1 page to what the reader
  is likely to be concerned is the strongest source of Noise in the
  observations does not seem unreasonable to us.

  We also think Section 4.3 and 4.4 are useful.  We are unaware of any
  predictions in the literature that make it easy for the reader to
  estimate the required duration of the observations describe in our
  manuscript.  Specifically, we know of no such comparison for state
  of the art facilities that is up to date for late 2015, nor do we
  know of any consolidation of this information in one publication.
  Predictions for future facilities become rapidly out of date as
  projects are descoped (or "rebaselined" in the case of the SKA), and
  many sensitivity predictions for existing instruments were made
  before the instrument was built.  Including the latest measured
  performance or design specifications is crucial to assess observing
  feasibility.  In section 4.4, we discuss the important point that
  one can exploit the unique time variability of these signatures for
  detection.  We think this is a crucial point to consider, as direct
  imaging would be impossible in the face of confusion from both
  Galactic and extragalactic sources (which are brighter but will not
  vary on the same time scale).

### 2015-11-12 ms_20150929submittedtoapj.tex
writing draft of reply to referee
adding theta = resolution to 4.3

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{deluxetable*}{cccccc}
\tabletypesize{\footnotesize}
\tablecolumns{6}
\tablecaption{Detectability of Hot Jupiters with Current 
and Future Radio Telescopes.\label{tab:sens}}
\tablehead{
Instrument & Band  & \sigRMS\tablenotemark{a} & $\tau$\tablenotemark{b}  & Resolution & $\sigma_c$\tablenotemark{c}\\
           & (MHz) & ($\mu$Jy bm$^{-1}$)         &  (hr)                    & (\arcsec)  & ($\mu$Jy bm$^{-1}$)} 
\startdata
LOFAR-LBA  &  30-80  & 1100  & 230000             & 10   & 57 \\ 
LOFAR-HBA  & 110-200 & 180   &   6500             & 4.1  & 1.1 \\  
HERA       & 50-250  &  76   & 1200               & 410  & 36000\\ 
GMRT       & 130-190 & 162   &   5400             & 16.5 & 57 \\ 
%LWA       &  30-80  & 50000 &  702127141         & 10.3 & 42.2 \\
%MWA       &  30-80  & 24000 &  176223211         & 1030 & 4.2$\times10^5$ \\
SKA-Low (part) & 50-200 & 5.9 &   7.2             & 7.6  & 9.7 \\
SKA-Low (full) & 50-350 & 1.8 &  0.7              & 4.8  & 1.5 \\
%VLA 4-Band& 30-80   & 3300  & 2312058            & 28.6 & 330 \\
VLA-LOBO   &  50-350 &  77   &  1200              & 8.6  & 10.4
\enddata
\tablenotetext{a}{RMS noise in a 1-hour observation, computed for imaging of the full band reported 
in this table (Equation~\ref{eq:rms_sens}).}
\tablenotetext{b}{The integration time in hours required to reach an RMS of 2~$\mu$Jy (i.e.\ 5-$\sigma$ detection
of a typical source at 100~pc with 10~$\mu$Jy bm$^{-1}$ average flux).}
\tablenotetext{c}{Source confusion RMS contribution.}
\end{deluxetable*}
%
% HERA
% http://reionization.org/wp-content/uploads/2015/05/HERA0_imaging.pdf
%
% LOFAR
% https://www.astron.nl/~heald/test/sens.php; 
% 24 core stations, 16 remote, 8 international, 488 subbands
%
% GMRT
% http://gmrt.ncra.tifr.res.in/~astrosupp/obs_setup/sensitivity.html
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

An additional factor that affects detectability is source confusion, the imaging noise associated with unresolved interloping radio sources in an observation.
Source confusion is characterized by the standard deviation $\sigma_c$ in the surface brightness of an image due to one or more unresolved sources in the beam solid angle (for a review, see \cite{Condon1974,Condon2012}).
To estimate $\sigma_c$, reported in Table~\ref{tab:sens}, we use the relations provided by \cite{condon+ransom2016}, reproduced below:
%%%
\begin{eqnarray}
\sigma_c & \approx & 0.2 \left( \frac{\nu}{1 \rm~GHz} \right)^{-0.7} \left( \frac{\theta}{1\arcmin} \right)^2\!\!\!\!,~(\theta > 10\arcsec) \\
\sigma_c & \approx & 2.2 \left( \frac{\nu}{1 \rm~GHz} \right)^{-0.7} \left( \frac{\theta}{1\arcmin} \right)^{10/3}\!\!\!\!\!\!\!\!\!\!,~(\theta < 10\arcsec).
\end{eqnarray}
%%%
As Table~\ref{tab:sens} illustrates, source confusion is likely to exceed the 2 $\mu$Jy detection limit of a typical RGHJ source for most
telescopes under consideration.  However, source confusion can be mitigated through differencing observations that cancel out background sources.
Potential axes for such cancellation are polarization and time.
The source confusion limit is dominated by active galactic nuclei, which have low polarization fractions $\lesssim2.5\%$ \citep{Stil2014} and slowly vary with time.
Therefore both the circularly polarized nature of the cyclotron emission from planets and the time variability (discussed in \S\ref{ss:timevariability}) partially mitigate the source confusion limit, allowing planetary radio emission, like other time-variable sources, to be detected below the imaging confusion limit.
We therefore expect confusion not to be a strong limitation to the detectability of radio emission from RGHJs.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{deluxetable*}{cccccc}
\tabletypesize{\footnotesize}
\tablecolumns{6}
\tablecaption{Detectability of Hot Jupiters with Current 
and Future Radio Telescopes.\label{tab:sens}}
\tablehead{
Instrument & Band  & \sigRMS\tablenotemark{a} & $\tau$\tablenotemark{b}  & $\theta$\tablenotemark{c}  & $\sigma_c$\tablenotemark{d}\\
           & (MHz) & ($\mu$Jy bm$^{-1}$)         &  (hr)                    & (\arcsec)  & ($\mu$Jy bm$^{-1}$)} 
\startdata
LOFAR-LBA  &  30-80  & 1100  & 230000             & 10   & 57 \\ 
LOFAR-HBA  & 110-200 & 180   &   6500             & 4.1  & 1.1 \\  
HERA       & 50-250  &  76   & 1200               & 410  & 36000\\ 
GMRT       & 130-190 & 162   &   5400             & 16.5 & 57 \\ 
%LWA       &  30-80  & 50000 &  702127141         & 10.3 & 42.2 \\
%MWA       &  30-80  & 24000 &  176223211         & 1030 & 4.2$\times10^5$ \\
SKA-Low (part) & 50-200 & 5.9 &   7.2             & 7.6  & 9.7 \\
SKA-Low (full) & 50-350 & 1.8 &  0.7              & 4.8  & 1.5 \\
%VLA 4-Band& 30-80   & 3300  & 2312058            & 28.6 & 330 \\
VLA-LOBO   &  50-350 &  77   &  1200              & 8.6  & 10.4
\enddata
\tablenotetext{a}{RMS noise in a 1-hour observation, computed for imaging of the full band reported 
in this table (Equation~\ref{eq:rms_sens}).}
\tablenotetext{b}{The integration time in hours required to reach an RMS of 2~$\mu$Jy (i.e.\ 5-$\sigma$ detection
of a typical source at 100~pc with 10~$\mu$Jy bm$^{-1}$ average flux).}
\tablenotetext{c}{Resolution of the instrument.}
\tablenotetext{d}{Source confusion RMS contribution.}
\end{deluxetable*}
%
% HERA
% http://reionization.org/wp-content/uploads/2015/05/HERA0_imaging.pdf
%
% LOFAR
% https://www.astron.nl/~heald/test/sens.php; 
% 24 core stations, 16 remote, 8 international, 488 subbands
%
% GMRT
% http://gmrt.ncra.tifr.res.in/~astrosupp/obs_setup/sensitivity.html
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

An additional factor that affects detectability is source confusion, the imaging noise associated with unresolved interloping radio sources in an observation.
Source confusion is characterized by the standard deviation $\sigma_c$ in the surface brightness of an image due to one or more unresolved sources in the beam solid angle, or resolution element, $\theta$ (for a review, see \cite{Condon1974,Condon2012}).
To estimate $\sigma_c$, reported in Table~\ref{tab:sens}, we use the relations provided by \cite{condon+ransom2016}, reproduced below:
%%%
\begin{eqnarray}
\sigma_c & \approx & 0.2 \left( \frac{\nu}{1 \rm~GHz} \right)^{-0.7} \left( \frac{\theta}{1\arcmin} \right)^2\!\!\!\!,~(\theta > 10\arcsec) \\
\sigma_c & \approx & 2.2 \left( \frac{\nu}{1 \rm~GHz} \right)^{-0.7} \left( \frac{\theta}{1\arcmin} \right)^{10/3}\!\!\!\!\!\!\!\!\!\!,~(\theta < 10\arcsec).
\end{eqnarray}
%%%
As Table~\ref{tab:sens} illustrates, source confusion is likely to exceed the 2 $\mu$Jy detection limit of a typical RGHJ source for most
telescopes under consideration.  However, source confusion can be mitigated through differencing observations that cancel out background sources.
Potential axes for such cancellation are polarization and time.
The source confusion limit is dominated by active galactic nuclei, which have low polarization fractions $\lesssim2.5\%$ \citep{Stil2014} and slowly vary with time.
Therefore both the circularly polarized nature of the cyclotron emission from planets and the time variability (discussed in \S\ref{ss:timevariability}) partially mitigate the source confusion limit, allowing planetary radio emission, like other time-variable sources, to be detected below the imaging confusion limit.
We therefore expect confusion not to be a strong limitation to the detectability of radio emission from RGHJs.
